---
title: "dsp4sh_ref_states_figs"
author: "Katy Dynarski"
date: "2024-08-07"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(here)
library(janitor)
library(aqp)
library(flextable)
library(lme4)
library(multcomp)
library(multcompView)
library(ggeffects)
library(ggrepel)
library(ggspatial)
library(ggeasy)
library(ggpubr)
library(ggtern)
library(png)
library(magick)
library(soiltexture)
library(paletteer)
library(cowplot)
library(tidyverse)

# Load data
soc_pedon <- read.csv(here("data_processed", "05_soc_pedon_clim.csv"))
soc_horizon <- read.csv(here("data_processed", "05_soc_horizon_clim.csv"))
surf <- read.csv(here("data_processed", "05_surface_horizons.csv"))
project <- read.csv(here("data_processed", "05_project_data.csv"))
eco_sites <- read.csv(here("data_processed", "05_eco_sites.csv"))
shape_all <- read.csv(here("data_processed", "shape_scores_all.csv"))
shape_spatial <- read.csv(here("data_processed", "shape_spatial.csv")) %>% clean_names()
ssurgo_texture <- read.csv(here("data_processed", "05_ssurgo_texture.csv"))
esm_all <- read.csv(here("data_processed","esm_all.csv"))

# Make list with mean and sd functions - useful for generating summary tables
mean_sd <- list(
  mean = ~round(mean(.x, na.rm = TRUE), 2), 
  sd = ~round(sd(.x, na.rm = TRUE), 2)
)

min_max <- list(
  min = ~min(.x, na.rm=TRUE), 
  max = ~max(.x, na.rm=TRUE)
)

# Labels
project_labels <- c("UnivOfMinnesota" = "University of Minnesota",
                    "UTRGV" = "University of Texas - RGV",
                    "WashingtonState" = "Washington State",
                    "TexasA&MPt-1" = "Texas A&M - 1",
                    "TexasA&MPt-2" = "Texas A&M - 2",
                    "NCState" = "North Carolina State",
                    "Illinois" = "University of Illinois",
                    "OregonState" = "Oregon State",
                    "KansasState" = "Kansas State",
                    "UConn" = "University of Connecticut")

project_labels2 <- c("UnivOfMinnesota" = "University \nof \nMinnesota",
                    "UTRGV" = "University \nof Texas\nRGV",
                    "WashingtonState" = "Washington\nState",
                    "TexasA&MPt-1" = "Texas \nA&M-1",
                    "TexasA&MPt-2" = "Texas \nA&M-2",
                    "NCState" = "North \nCarolina \nState",
                    "Illinois" = "University \nof \nIllinois",
                    "OregonState" = "Oregon\nState",
                    "KansasState" = "Kansas\nState",
                    "UConn" = "University \nof \nConnecticut")

# Themes
theme_katy <- function(base_size=14) {
  theme_classic(base_size=base_size) %+replace%
    theme(# Legend
      legend.title=element_text(size=rel(1)), legend.text=element_text(size=rel(.8)),
      #axes                      
      axis.text=element_text(size=rel(.8)),
      axis.title=element_text(size=rel(1)))
}

theme_katy_grid <- function(base_size=14) {
  theme_classic(base_size=base_size) %+replace%
    theme(# Legend
      legend.title=element_text(size=rel(1)), legend.text=element_text(size=rel(.8)),
      #axes                      
      axis.text=element_text(size=rel(.8)),
      axis.title=element_text(size=rel(1)),
      # set margins (necessary for figure to not awkwardly overlap when using plot_grid in cowplot)
      plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"))
}
```

# Overview of DSP4SH Climate and Soil Texture

## Figure 1A - Dual-axis figure showing mean annual temperature and precipitation for each site

```{r figure_1a}
# Compile data
clim <- soc_pedon %>%
  distinct(project, dsp_plot_id, label, mat, map) %>%
  group_by(project) %>%
  summarize(mean_mat = mean(mat),
            mean_map = mean(map))

# Make the plot
proj_tem_prep_clim <- ggplot(clim, aes(x = project)) +
  # Add temperature points (scaled for dual y-axis)
  geom_point(aes(y = mean_mat * 40), color = "#CC5800FF", size = 4, shape = 19) + # Adjust scale factor as needed
  # Add precipitation points
  geom_point(aes(y = mean_map), color = "#1E8E99FF", size = 4, shape = 17) +
  scale_x_discrete(labels=project_labels, name="Project") + 
  # Customize y-axes
  scale_y_continuous(
    name = "Mean Annual Precipitation (mm)",
    sec.axis = sec_axis(~ . / 40, name = "Mean Annual Temperature (°C)")) + # Adjust scale factor
  theme_katy(base_size=12) +
  theme(axis.title.y=element_text(color="#1E8E99FF"),
        axis.title.y.right=element_text(color="#CC5800FF"),
        axis.text.x=element_text(angle=45, hjust=1),
        plot.margin = margin(l=30))

proj_tem_prep_clim
```

## Figure 1B - Soil texture triangle showing mean SSURGO texture in top horizon for each project

Make the USDA soil texture triangle to overlay:

```{r usda_texture_triangle}

# USDA dataset comes pre-loaded in ggtern package
data(USDA)

# rename the labels to be shorter for easier printing
USDA <- USDA %>% 
  rename("clay_pct" = Clay, "sand_pct" = Sand, "silt_pct" = Silt) %>% 
  mutate(
    label = case_when(
      Label == "Clay" ~ "C",
      Label == "Sandy Clay" ~ "SC",
      Label == "Sandy Clay Loam" ~ "SCL",
      Label == "Sandy Loam" ~ "SL",
      Label == "Loamy Sand" ~ "LS",
      Label == "Sand" ~ "S",
      Label == "Clay Loam" ~ "CL",
      Label == "Silt Loam" ~ "SiL",
      Label == "Silty Clay" ~ "SiC",
      Label == "Silty Clay Loam" ~ "SiCL",
      Label == "Silt" ~ "Si",
      Label == "Loam" ~ "L",
      TRUE ~ NA_character_
    ))
USDA_text <- USDA  %>% 
  group_by(label) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
USDA_text
```

Now, plot the SSURGO data on the textural triangle:

```{r fig_1b}
# Make a new column with both project and soil name for plotting
project_labels_df <- data.frame(project_labels) %>%
  rownames_to_column() %>%
  rename(project = rowname)

ssurgo_texture2 <- ssurgo_texture %>%
  left_join(project_labels_df, by="project") %>%
  unite("project_soil", c("project_labels", "soil"), remove=FALSE, sep=" - ")

# Plot texture with SSURGO data only, point fill color by project/soil combo
# overlay textural triangle
theme_set(theme_bw())
triangle_plot <- ggtern(ssurgo_texture2, aes(x=sand_pct, y=clay_pct, z=silt_pct, 
                                          color=project_soil)) +
  geom_polygon(data=USDA,
               aes(fill = label),
               alpha = 0.0,
               linewidth = 0.5,
               color = "black",
               show.legend = FALSE) +
  geom_text(data = USDA_text,
            aes(label = label),
            color = 'grey40',
            fontface = "bold",
            size = 5) +
  geom_point(size=4) +
  theme_showarrows() +
  labs(yarrow = "Clay (%)",
       zarrow = "Silt (%)",
       xarrow = "Sand(%)",
       x="", y="", z="",
       color = "Project")  +
  theme_clockwise() +
  scale_color_paletteer_d("rcartocolor::Safe", name="Project & Soil Series")
triangle_plot
```

```{r Figure 1}
# Put two panels together

# ggtern does not play well with cowplot - so call in the texture triangle figure as a PNG file and add to a grid with a blank space
tt_file <- readPNG(here("figs", "ref_states_figs", "fig1b_ssurgo_texture.png"))

fig1_grid_space <- plot_grid(proj_tem_prep_clim, NULL,
                          nrow=2, rel_heights=c(1,1),
                          labels=c("A", "B"))
fig1 <- ggdraw(fig1_grid_space) +
  draw_image(tt_file, x=0, y=0, vjust=0.23)

fig1
```

# Summary Figures of SOC Stocks

## Summary table of SOC stocks across projects, soils, and treatments

Mean, standard deviation, and n for SOC stocks calculated to 30 cm and
100 cm for all soils in DSP4SH projects.

```{r table_1}
soc_summary <- soc_pedon %>%
  group_by(project, soil, label) %>%
  summarize(across(soc_stock_0_30cm:soc_stock_100cm, mean_sd), n=n())
flextable(soc_summary)

# range in SOC values
soc_min_max <- soc_pedon %>%
  summarize(across(soc_stock_0_30cm:soc_stock_100cm, min_max), n=n())
flextable(soc_min_max)

# which is the min?
soc_pedon %>% slice_min(soc_stock_100cm, n = 1) %>%
  select(project, label, dsp_pedon_id, soc_stock_100cm)

# which is the max?
soc_pedon %>% slice_max(soc_stock_100cm, n = 1) %>%
  select(project, label, dsp_pedon_id, soc_stock_100cm)
```

## Fig 2 - Boxplots of total SOC stocks under different management treatments

Plot boxplots of total SOC stocks under different management treatments:

```{r fig_2}
soc_pedon_toplot <- soc_pedon %>%
  filter(project!="TexasA&MPt-2") %>% # Exclude Texas A&M pt 2 because data was only collected to ~20 cm, so no stocks were calculated
  mutate(label=factor(label, levels=c("BAU", "SHM", "Ref")))

# Boxplot comparing total SOC stocks (100 cm) between treatments within soil types
soc100_boxplot <- ggplot(soc_pedon_toplot, 
                         aes(x=project, y=soc_stock_100cm, fill=label)) +
  geom_boxplot() +
  labs(x="Project", y="SOC stock to 100 cm depth (Mg/ha)") +
  scale_x_discrete(labels=project_labels) +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy_grid() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# Boxplot comparing total SOC stocks (30 cm) between treatments within soil types
soc30_boxplot <- ggplot(soc_pedon_toplot, 
                         aes(x=project, y=soc_stock_0_30cm, fill=label)) +
  geom_boxplot() +
  labs(x="Project", y="SOC stock to 30 cm depth (Mg/ha)") +
  scale_x_discrete(labels=project_labels) +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy_grid() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

soc_grid <- plot_grid(soc30_boxplot + 
                               theme(legend.position="none", axis.title.x=element_blank(), axis.text.x=element_blank()),
          soc100_boxplot + theme(legend.position="none"),
          ncol=1, labels=c("A", "B"), rel_heights=c(1, 1.4))
soc_leg <- get_legend(soc100_boxplot)

plot_grid(soc_grid, soc_leg, rel_widths = c(1, .23))
```

Takeaways:

-   Significant variability in SOC stocks between sites (expected)

-   Differences between treatments more apparent at 30 cm depth than 100
    cm depth

-   Reference sites are significantly higher in C at some, but not all
    sites (Illinois, UConn, University of Minnesota, UTRGV)

-   Differences between SHM and BAU are infrequently observed

## Boxplots of SOC stock data calculated via ESM

```{r fig_s1}
# ESM data contains many different ways to calculate - select only the SOC stocks calculated via standard depth increments, individual projects as reference soils, and using the minimum soil mass as a reference
esm_standard_min <- esm_all %>%
  filter(depth_increments == "standard", ref_data =="indv_project",
         method_long== "esm2_min")

# Calculate total SOC stocks
esm_standard_min_totals <- esm_standard_min %>%
  group_by(project, label, dsp_pedon_id, depth_increments) %>%
  summarize(soc_0to30 = sum(soc[apparent_depth=="0-10 cm" | apparent_depth=="10-30 cm"]),
            soc_0to100 = sum(soc))

# Boxplot comparing total SOC stocks (30 cm) between treatments within soil types
soc30_esm_boxplot <- ggplot(esm_standard_min_totals, 
                         aes(x=project, y=soc_0to30, fill=factor(label, levels=c("BAU", "SHM", "Ref")))) +
  geom_boxplot() +
  labs(x="Project", y="SOC stock to 30 cm depth (Mg/ha)") +
  scale_x_discrete(labels=project_labels) +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy_grid() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# Boxplot comparing total SOC stocks (100 cm) between treatments within soil types
soc100_esm_boxplot <- ggplot(esm_standard_min_totals, 
                         aes(x=project, y=soc_0to100, fill=factor(label, levels=c("BAU", "SHM", "Ref")))) +
  geom_boxplot() +
  labs(x="Project", y="SOC stock to 100 cm depth (Mg/ha)") +
  scale_x_discrete(labels=project_labels) +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy_grid() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

soc_esm_grid <- plot_grid(soc30_esm_boxplot + theme(legend.position="none", axis.title.x=element_blank(), axis.text.x=element_blank()),
          soc100_esm_boxplot + theme(legend.position="none"),
          ncol=1, labels=c("A", "B"), rel_heights=c(1, 1.4))
soc_esm_leg <- get_legend(soc100_esm_boxplot)

plot_grid(soc_esm_grid, soc_esm_leg, rel_widths = c(1, .23))
```

## Comparison of ESM results to fixed depth results

```{r fig_s2}
esm_standard_min_totals_long <- esm_standard_min_totals %>%
  pivot_longer(soc_0to30:soc_0to100, names_to="depth", values_to="soc_stock") %>%
  select(-depth_increments) %>%
  mutate(depth = case_when(depth== "soc_0to30" ~ "0to30cm",
                               depth=="soc_0to100" ~ "0to100cm"))

soc_pedon_fixed <- soc_pedon %>%
  select(project, label, dsp_pedon_id, soc_stock_0_30cm, soc_stock_100cm) %>%
  pivot_longer(soc_stock_0_30cm:soc_stock_100cm,
               names_to="depth", values_to="soc_stock") %>%
    mutate(depth = case_when(depth== "soc_stock_0_30cm" ~ "0to30cm",
                               depth=="soc_stock_100cm" ~ "0to100cm"))

soc_fixed_esm <- soc_pedon_fixed %>%
  left_join(esm_standard_min_totals_long, by=c("project", "label", "dsp_pedon_id","depth"), suffix=c("_fixed", "_esm")) %>%
  pivot_longer(soc_stock_fixed:soc_stock_esm, names_to="calc_method", values_to="soc_stock") %>%
  mutate(depth = factor(depth, levels=c("0to30cm", "0to100cm"))) %>%
  filter(project!="TexasA&MPt-2")

# Make plot
ggplot(soc_fixed_esm, aes(x=label, y=soc_stock, fill=calc_method)) +
  geom_boxplot() +
  facet_grid(project ~ depth, scales="free_y", labeller=labeller(project = project_labels2, 
                                                                 depth=c("0to100cm"="0 to 100 cm", "0to30cm"="0 to 30 cm"))) +
  labs(x="Management", y="SOC Stock (Mg/ha)") +
  scale_fill_paletteer_d("nationalparkcolors::Arches", name="Calculation method",
                         labels=c("ESM (Cubic Spline)", "Fixed Depth")) +
  theme_katy()
```

Table of difference in SOC stocks calculated via fixed depth vs. ESM:

```{r table_s1}
# Calculate differences
soc_stock_comparison <- soc_fixed_esm %>%
  pivot_wider(names_from="calc_method", values_from="soc_stock") %>%
  group_by(project, label, depth) %>%
  summarize(across(soc_stock_fixed:soc_stock_esm, ~round(mean(.x, na.rm=TRUE),1))) %>%
  mutate(difference = round(soc_stock_fixed - soc_stock_esm,1)) %>%
  mutate(percent_diff = round((difference / soc_stock_esm)*100,1)) %>%
  mutate(label=factor(label, levels=c("BAU", "SHM", "Ref"))) %>%
  arrange(project, label, depth)

flextable(soc_stock_comparison)

soc_stock_comparison_summary <- soc_stock_comparison %>%
  ungroup() %>%
  group_by(depth) %>%
  summarize(mean_diff = round(mean(difference),1),
            mean_percent_diff = round(mean(percent_diff),1))
flextable(soc_stock_comparison_summary)
```

Mean difference in SOC stocks between fixed depth and ESM is 9.5 Mg/ha (greater in fixed depth vs ESM), mean percent difference is 14.3%.

## Fig 3 - Ribbon plot of SOC stocks by depth (continuous):

```{r fig_3}
# Promote horizon data to SPC
soc_spc <- soc_horizon
depths(soc_spc) <- dsp_pedon_id ~ hrzdep_t + hrzdep_b
hzdesgnname(soc_spc) <- 'hzdesg'
# promote project and label to site-level so they can be used as grouping variables
site(soc_spc) <- ~ project + label + soil

# Calculate stocks by depth increment for each project and management condition
slab_ref <- aqp::slab(subset(soc_spc, label=="Ref"),
                      fm = project ~ soc_stock_hrz,
                      slab.structure = seq(0,100,by=10)) %>%
  mutate(label="Ref")

slab_shm <- aqp::slab(subset(soc_spc, label=="SHM"),
                      fm = project ~ soc_stock_hrz,
                      slab.structure = seq(0,100,by=10)) %>%
  mutate(label="SHM")
slab_bau <- aqp::slab(subset(soc_spc, label=="BAU"),
                      fm = project ~ soc_stock_hrz,
                      slab.structure = seq(0,100,by=10)) %>%
  mutate(label="BAU")

# Put management conditions together
slab_mgmt <- bind_rows(slab_ref, slab_shm, slab_bau)

# Plot with all mgmt together
ggplot(slab_mgmt, aes(x=top, y=p.q50)) +
  geom_line(linewidth=1.2, aes(color=label)) +
  geom_ribbon(aes(ymin=p.q25, ymax=p.q75, x=top, fill=label), alpha=0.2) +
  xlim(c(100,0)) +
  coord_flip() +
  labs(title="SOC Stocks by Depth", x="Depth (cm)", y="SOC (Mg/ha)") +
  facet_wrap(~ project, scales = "free_x", labeller=labeller(project=project_labels)) +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                    guide="none") +
  scale_color_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                     breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy()
```

Takeaways:

-   Very few projects actually showed greater SOC stocks throughout the
    soil profile (Minnesota, Texas A&M, UTRGV)
-   Differences in C stocks tended to be most apparent at shallow depths
    (though see Texas A&M Pt 1 and Minnesota for exceptions where
    differences were greater deeper in the soil profile)

# Test effect of treatment on SOC stocks and concentrations with mixed linear model

## Effect of treatment on SOC stocks to 100 cm depth

```{r mixed linear model of SOC stocks to 100 cm by label and soil}
# Remove NAs from data
soc100_clean <- soc_pedon %>%
  select(dsp_pedon_id, project, label, soil, soc_stock_100cm) %>%
  na.omit() 

# Random effects: project
# Fixed effects: label
soc_stock100_mixed <- lmer(soc_stock_100cm ~ label + (1|project), data = soc100_clean)
summary(soc_stock100_mixed)

# Test significance of treatment by comparing full and reduced models, use likelihood ratio test
stock100_mixed_full <- lmer(soc_stock_100cm ~ label + (1|project), data = soc100_clean, REML = FALSE)
stock100_mixed_reduced <- lmer(soc_stock_100cm ~ (1|project), data = soc100_clean, REML = FALSE)
anova(stock100_mixed_full, stock100_mixed_reduced, text="Chisq")

# Tukey post-hoc
soc_stock100_mixed_tukey <- glht(soc_stock100_mixed, linfct = mcp(label = 'Tukey'))
summary(soc_stock100_mixed_tukey)
```

**Results:** Full model is significantly different from reduced model -
there is a significant effect of treatment when between-project
variation is controlled for. Tukey HSD post-hoc test shows that Ref SOC
stock is significantly different from BAU and Ref groups.

## Effect of treatment on SOC stocks to 30 cm depth

```{r mixed linear model of SOC stocks to 30 cm by label and soil}
# Remove NAs from data
soc30_clean <- soc_pedon %>%
  select(dsp_pedon_id, project, label, soil, soc_stock_0_30cm) %>%
  na.omit() 

# Random effects: project
# Fixed effects: label
soc_stock30_mixed <- lmer(soc_stock_0_30cm ~ label + (1|project), data = soc30_clean)
summary(soc_stock30_mixed)

# Test significance of treatment by comparing full and reduced models, use likelihood ratio test
stock30_mixed_full <- lmer(soc_stock_0_30cm ~ label + (1|project), data = soc30_clean, REML = FALSE)
stock30_mixed_reduced <- lmer(soc_stock_0_30cm ~ (1|project), data = soc30_clean, REML = FALSE)
anova(stock30_mixed_full, stock30_mixed_reduced, text="Chisq")
# Full model is significantly different from reduced model - there is a significant effect of treatment when between-project variation is controlled for 

# Tukey post-hoc
soc_stock30_mixed_tukey <- glht(soc_stock30_mixed, linfct = mcp(label = 'Tukey'))
summary(soc_stock30_mixed_tukey)
```

**Results:** Full model is significantly different from reduced model -
there is a significant effect of treatment when between-project
variation is controlled for. Tukey HSD post-hoc test shows that Ref SOC
stock is significantly different from BAU and Ref groups.

## Plot results of mixed linear model

```{r soc_lmer_plot}
# Plot predictions of 30cm SOC stocks under each treatment
pred_stock30 <- ggpredict(soc_stock30_mixed, terms = c("label"))
pred30 <- ggplot(pred_stock30, aes(x=x, y=predicted)) +
  geom_point() +
  geom_errorbar(aes(x=x, ymin=conf.low, ymax=conf.high)) +
  labs(x="Management", y="Predicted SOC stock to 30 cm (Mg/ha)") +
  theme_katy_grid()

# Plot predictions of 100cm SOC stocks under each treatment
pred_stock100 <- ggpredict(soc_stock100_mixed, terms = c("label"))

pred100 <- ggplot(pred_stock100, aes(x=x, y=predicted)) +
  geom_point() +
  geom_errorbar(aes(x=x, ymin=conf.low, ymax=conf.high)) +
  labs(x="Management", y="Predicted SOC stock to 100 cm (Mg/ha)") +
  theme_katy_grid()

plot_grid(pred30, pred100,
          ncol=1, labels=c("A", "B"))
```

Takeaways:

-   Though not many significant differences are detected between
    treatments within each project, mixed linear model suggests that
    when accounting for between-project variability, SOC stocks (both
    30cm and 100cm depth) are significantly higher in the Ref condition
    vs. SHM and BAU.

-   Supports choice of Ref conditions as a broad concept.

# SHAPE Score Analysis

Plot SOC stocks vs SHAPE scores:

```{r plot SHAPE scores vs carbon stocks}
# Need to join SHAPE scores to SOC stock data
shape_soc_stock <- shape_spatial %>%
  select(dsp_pedon_id, score_mean_soc, gt_90_soc) %>%
  left_join(select(soc_pedon, dsp_pedon_id, project, soil, label, trt, lu, till, soc_stock_0_30cm, soc_stock_100cm),
            by="dsp_pedon_id")

# Details of logarithmic regression
shape_soc_lm <- lm(score_mean_soc ~ log(soc_stock_100cm), data=shape_soc_stock)
summary(shape_soc_lm)

# Plot
shape_soc_plot <- ggplot(shape_soc_stock, aes(x=soc_stock_100cm, y=score_mean_soc)) +
  geom_point(aes(colour=label)) + 
  geom_smooth(method="lm", formula=y~log(x), color="black") +
  annotate(geom="text", x=250, y=0.22,
           label=expression(atop("y = -0.90 + 0.32*log(x)", R^2~"= 0.45, p < 0.001"))) +
  labs(x="SOC stock to 100 cm depth (Mg/ha)",
       y="SHAPE Score (Peer Group Percentile)") +
  scale_colour_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                    breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy()
```

Boxplots of SHAPE SOC scores:

```{r fig_4}
# Join in project data
shape_spatial_proj <- shape_spatial %>%
  select(dsp_pedon_id, score_mean_soc, gt_90_soc) %>%
  left_join(select(project, dsp_plot_id, dsp_pedon_id, project, soil, label, trt, lu, till), by="dsp_pedon_id") %>%
  mutate(shape_source = "spatial")

# Boxplot showing SHAPE scores in different management categories, separated by project
shape_boxplot_all <- ggplot(shape_spatial_proj, aes(x=project, y=score_mean_soc, fill=label)) +
  geom_boxplot() +
  geom_abline(intercept=0.90, slope=0, linetype="dashed") +
  geom_abline(intercept=0.75, slope=0, linetype="dashed") +
  labs(x="Project", y="SHAPE Score (Peer Group Percentile)") +
  scale_x_discrete(labels=project_labels) +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                    breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy() +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# Boxplot showing SHAPE scores in different management categories (no separation by project)
shape_boxplot_condensed <- ggplot(shape_spatial_proj, aes(x=factor(label, levels=c("BAU", "SHM", "Ref")), 
                                                          y=score_mean_soc, 
                      fill=factor(label, levels=c("BAU", "SHM", "Ref")))) +
  geom_boxplot() +
  geom_abline(intercept=0.90, slope=0, linetype="dashed") +
  geom_abline(intercept=0.75, slope=0, linetype="dashed") +
  labs(y="SHAPE Score (Peer Group Percentile)",
       x="Management System") +
  scale_fill_manual(values=c("#FED789FF","#72874EFF","#476F84FF"),
                    breaks=c("BAU", "SHM", "Ref"), 
                    name="Management") +
  theme_katy() +
  theme(plot.title=element_text(hjust=0.5), legend.position="none")

# Panel figure of all SHAPE and consolidated SHAPE AND SOC stock
shape_grid_bottom <- plot_grid(shape_boxplot_condensed + theme(legend.position="none"),
          shape_soc_plot + theme(legend.position="none"),
          ncol=2, labels=c("B", "C"))

shape_grid <- plot_grid(shape_boxplot_all + theme(legend.position="none"),
                        shape_grid_bottom,
                        ncol=1, labels=c("A", ""), rel_heights=c(1.23, 1))

shape_leg <- get_legend(shape_boxplot_all)

plot_grid(shape_grid, shape_leg, rel_widths = c(1, 0.23))
```

Overall, the SHAPE scores support the Ref/SHM/BAU groupings:

-   Very few projects had median Ref SHAPE score at 90th percentile or
    above (Illinois, Oregon State, University of Minnesota, and UTRGV)

-   Most projects had median Ref SHAPE score at 75th percentile or above
    (only Texas A&M projects do not hit 75th percentile - wonder if this
    is a spatial issue)

-   Only one project has SHM SHAPE score at 90th percentile (UConn)

-   Few projects had SHM SHAPE scores at 75th percentile (University of
    Minnesota, Oregon State, NC State is just a hair below the 75th
    percentile)

-   No projects had BAU SHAPE scores at or above the 75th percentile

## SHAPE Scores and ESD-STMs
```{r join ecological site data to SHAPE data}
eco_sites_trt <- eco_sites %>%
  left_join(select(project, dsp_plot_id, label, trt), by="dsp_plot_id") %>%
  distinct()

eco_sites_shape <- eco_sites %>%
  left_join(shape_spatial_proj, by="dsp_plot_id")
```

Plot SHAPE scores and SOC% corresponding to different ecological states for project that correlated sampling sites with ecological states - UTRGV is the only project that did this.

```{r fig_5}
# Compile data - need surface SOC%
utrgv_soc_surf <- surf %>%
  select(project, dsp_pedon_id, soc_pct) %>%
  filter(project=="UTRGV")

shape_soc_utrgv <- eco_sites_shape %>%
  filter(project.y=="UTRGV") %>%
  select(project.y, dsp_pedon_id, ecological_site, ecological_state, soil, label, score_mean_soc) %>%
  rename(project = project.y,
         soc_shape = score_mean_soc) %>%
  left_join(utrgv_soc_surf, by=c("dsp_pedon_id", "project")) %>%
  rename(soc_measured = soc_pct) %>%
  pivot_longer(cols=soc_shape:soc_measured,
               names_to="type", values_to="value")

# Make plot
ggplot(shape_soc_utrgv, aes(x=ecological_state, 
                          y=value, 
                          fill=ecological_state)) +
  geom_boxplot() +
  labs(x="Ecological state",
       title="UTRGV - Loamy Bottomland Ecological Site",
       y=NULL) +
  facet_wrap(vars(type), scales="free",
             strip.position = "left", 
             labeller = as_labeller(c(soc_measured = "Measured SOC (%)", soc_shape = "SOC SHAPE Score (Peer Group Percentile"))) +
  scale_fill_manual(values=c("#FED789FF", "#476F84FF")) +
  theme_katy() +
  theme(legend.position="none",
        strip.background = element_blank(),
        strip.placement = "outside")
```

The tricky thing here is that woodland state is not actually the reference state for this ecological site - the reference state is savannah!